Looking for CDI Bias: WG Comprehension

### 2021-07-30 01-preliminary-WGprod.Rmd
multipleGroup model comparisons

>>>>

# Get DIF Statistics

```{r, eval=F, include=F}
# which.par = c('a1','d')
# items2test = 1  # specify which items to test -- can split up 
# scheme = 'add' # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.
# mirtCluster() # allows parallel running of models
start_time = Sys.time()
d_s <- DIF(mg2, which.par='d', items2test = 1:2, scheme='drop')
end_time = Sys.time()
```

### 2021-07-30 02-preliminary-WGcomp.Rmd
multipleGroup model comparisons

# mother's years of education = proxy for SES
table(d_demo$mom_ed)
# skewed: only 72 with Primary/Some Secondary (most have Some College +)

@@

table(en_wg$category)
```

@@

mg1 <- multipleGroup(d_mat, 1, group = sex_group, SE = T) 
# Error: Multiple Group model will not be identified without proper constraints (groups contain missing data patterns where item responses have been completely omitted or, alternatively, the number of categories within each group is not equal to the total number of categories)
mg2 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)

m_sc2 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                       invariance=c('slopes', 'intercepts', 'free_var','free_means'))
mod_sc1 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000), #fixed means
                             invariance=c('slopes', 'intercepts', 'free_var'))
mod_fullconstrain <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts'))

@@

save(file = here("data/eng_wg_comp_sex_sexDIF.Rds"), 
     mg2, m_sc2, mod_sc1, mod_fullconstrain)
```

@@

mg2_df = extract_group_df(mg2)

@@

@@

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg2_df$d_diff_abs) + sd(mg2_df$d_diff_abs)
big_dif = mg2_df[which(mg2_df$d_diff_abs > thresh),]

@@

```{r, fig.width=8, fig.height=6}
ggplot(mg2_df, 
       aes(x = d_m, y = d_f)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

@@

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg2 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg2, method = "MAP")[,1])

@@

```{r}
d_demo %>% group_by(sex) %>% 
  summarise(ability_2pl = mean(ability),
            ability_sex = mean(ability_sex)) %>%
  kable(digits=2)
```

@@

@@

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg2, m_sc2, mod_sc1, mod_fullconstrain)
```

```{r}
load(here("data/eng_wg_comp_SES_DIF.Rds"))

@@

mg2_df = extract_ses_group_df(mg2)
```

@@

```{r, fig.width=8, fig.height=6}
ggplot(mg2_df, 
       aes(x = d_lo, y = d_hi)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for low-SES") + 
  ylab("Item Ease for high-SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

>>>>

# mother's years of education = proxy for SES
table(d_demo$mom_ed) %>% kable()
# skewed: only 72 with Primary/Some Secondary (most have Some College +)

@@

table(en_wg$category) %>% kable()
```

@@

mg1 <- multipleGroup(d_mat, 1, group = sex_group) # this example should work, according to docs
mg1 <- multipleGroup(d_mat, 1, group = sex_group, 
                     invariance = c("free_var", "free_means")) 
# Error: Multiple Group model will not be identified without proper constraints (groups contain missing data patterns where item responses have been completely omitted or, alternatively, the number of categories within each group is not equal to the total number of categories)

# unless there are possibly items

mg2 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)
#DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.

m_sc2 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                       invariance=c('slopes', 'intercepts', 'free_var','free_means'))

# constrain slopes, freely estimate latent group means (reference group constrained to a vector of 0)
mg3 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance =c('slopes', 'free_means'), verbose = T)

# constrain slopes, allow varying group means and variances -- very slow!
#mg4 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
#                     invariance =c('slopes','free_means', 'free_var'), verbose = T)

# fixed means, varying group variance
mod_sc1 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts', 'free_var'))
# fully-constrained
mod_fullconstrain <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts'))

# constrain intercepts to be equal across groups, but allow group slopes to vary
m_int <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                     invariance = 'intercepts')

anova(mod_sc1, mod_fullconstrain) # mod_fullconstrain preferred by AIC/BIC
anova(m_sc2, mod_fullconstrain) # mod_fullconstrain preferred by AIC/BIC

anova(mg2, mod_fullconstrain) # AIC prefers slope-invariant model; BIC prefers fully-constrained model

anova(m_int, mod_fullconstrain) # AIC prefers intercept-constrained model; BIC prefers fully-constrained

anova(mg3, mg2) # AIC//BIC prefer model with varying group means
#anova(mg3, mg4)

coef(mg3)[[2]]$GroupPars # Female intercept = 0.22 (Male = 0)

@@

save(file = here("data/eng_wg_comp_sex_sexDIF.Rds"), 
     mg2, mg3,  m_sc2, mod_sc1, mod_fullconstrain) # not run: mg4
```

@@

# slope-invariant, varying difficulty:
mg2_df = extract_group_df(mg2)
mg3_df = extract_group_df(mg3)

# same item estimates with/without free_means? 
cor(mg2_df$d_f, mg3_df$d_f) # .9999
cor(mg2_df$d_m, mg3_df$d_m) # .9998

@@

# fully-constrained is just the same as ordinary 2PL model, no? but we can use this model for DIF
mod_fullconstrain_df = extract_group_df(mod_fullconstrain)
# DIF(mod_fullconstrain, which.par = 'd')

#mod_int_df = extract_group_df(m_int)

mod_sc2_df = extract_group_df(m_sc2)

@@

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg3_df$d_diff_abs) + sd(mg3_df$d_diff_abs)
big_dif = mg3_df[which(mg3_df$d_diff_abs > thresh),]

@@

```{r, fig.width=8, fig.height=6}
ggplot(mg3_df, 
       aes(x = d_m, y = d_f)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

@@

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg3 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg3, method = "MAP")[,1])

fscores_mg2 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg2, method = "MAP")[,1])

# very similar
#plot(fscores_mg3$ability_sex, fscores_mg2$ability_sex, pch='.')

@@

```{r}
d_demo %>% group_by(sex) %>% 
  summarise(ability_2pl = mean(ability),
            ability_sex = mean(ability_sex)) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:30%;'")
```

@@

mg3 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = c('slopes', 'free_means'), verbose = T)

anova(mg2, mg3) # AIC/BIC prefer model with varying group mean

@@

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg2, mg3, m_sc2, mod_sc1, mod_fullconstrain)
```

```{r}
load(here("data/eng_wg_comp_SES_DIF.Rds"))
load(here("data/eng_wg_comp_mod_2pl.Rds"))

@@

mg2_df = extract_ses_group_df(mg2)
mg3_df = extract_ses_group_df(mg3)

coef(mg3)[[2]]$GroupPars # low-SES group (high is ref) intercept = .24
```

@@

```{r, fig.width=8, fig.height=6}
ggplot(mg3_df, 
       aes(x = d_lo, y = d_hi)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg3_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for low-SES") + 
  ylab("Item Ease for high-SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

What are the ability distributions by SES in the 2PL model?

d_demo %>% ggplot(aes(x=ability, fill=ses_group)) + 
  geom_density(alpha=.3) + 
  theme_classic()
```

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg2 <- tibble(data_id = rownames(d_mat[which(!is.na(d_demo$ses_group)),]), 
                             ability_ses = fscores(mg2, method = "MAP")[,1])

d_demo_ses <- d_demo %>% 
  filter(!is.na(ses_group)) %>% 
  left_join(fscores_mg2 %>% mutate(data_id = as.numeric(data_id)))

ggpairs(d_demo_ses, columns = c("age","comprehension","ability","ability_ses"),
        ggplot2::aes(colour=ses_group, alpha=.5)) + theme_classic()
```

```{r, echo=F}
d_demo_ses %>% group_by(ses_group) %>% 
  summarise(ability_2pl = mean(ability),
            ability_ses = mean(ability_ses)) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:30%;'")
```
